---
title: "EuroLeague Basketball"
author: "Jo Hardin"
format: 
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(DT)
library(fontawesome)
fontawesome::fa_html_dependency()
```

This week we’re exploring **EuroLeague Basketball**, the premier men’s club basketball competition in Europe.  

The dataset contains information on EuroLeague teams, including their country, home city, arena, seating capacity, and historical performance (Final Four appearances and titles won).  

The dataset is curated from publicly available sources such as Wikipedia and official EuroLeague records, and was packaged in the [EuroleagueBasketball R package](https://github.com/natanast/EuroleagueBasketball), with documentation available at [natanast.github.io/EuroleagueBasketball](https://natanast.github.io/EuroleagueBasketball/).

> "The EuroLeague is the top-tier European professional basketball club competition, widely regarded as the most prestigious competition in European basketball." — EuroLeague


## The Data

```{r}
euroleague_basketball <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-10-07/euroleague_basketball.csv') |> 
  mutate(capacity = str_split(Capacity, ", ")) |> 
  unnest(capacity) |> 
  mutate(capacity = as.numeric(str_replace_all(capacity, ",", ""))) |> 
  mutate(lastrank = as.numeric(str_extract(`Last season`, "\\d+")),
         lastrank = ifelse(is.na(lastrank), 20, lastrank)) |> 
  select(Team, Country, capacity, lastrank, FinalFour_Appearances, Titles_Won, everything())
```



##

```{r}
library(GGally)

ggparcoord(data = euroleague_basketball, 
           columns = 3:6,
           alphaLines = 1,
           groupColumn = 2,
           scale = "uniminmax") + 
  scale_x_discrete(labels = colnames(euroleague_basketball)[3:6])
```




```{r}
#| fig.alt: 'Parallel coordinate plot comparing EuroLeague basketball teams. The variables shown are team, country, stadium capacity, 2024-2025 rank, number of final four appearances, and number of titles won. There are a few Greek teams and Spanish teams which seem to have dominated the league in the past few decades.'
#| fig.cap: 'The two Greek teams Panathinaikos and Olympiacos seem to dominate the basketball EuroLeauge. Panathinaikos and Dubai Basketball have the biggest stadiums.'

library(plotly)

euro_plot_data <- euroleague_basketball[,1:6] |> 
  mutate(country_id = as.numeric(factor(Country)),
         team_id = as.numeric(factor(Team))) 


euro_plot_data |> 
plot_ly(
  type = 'parcoords',
  line = list(color = ~country_id,
              colorscale = 'Rainbow'),
  dimensions = list(
    list(label = "Team", values = ~team_id,
         tickvals = unique(euro_plot_data$team_id),
         ticktext = unique(euro_plot_data$Team)),
    list(label = "Country", values = ~country_id,
         tickvals = unique(euro_plot_data$country_id),
         ticktext = unique(euro_plot_data$Country)),
    list(label = "Capacity",  values = ~capacity),
    list(label = "2024-2025 Rank", values = ~lastrank,
         range = c(max(euro_plot_data$lastrank, na.rm = TRUE),
                   min(euro_plot_data$lastrank, na.rm = TRUE))),
    list(label = "FinalFour", values = ~FinalFour_Appearances),
    list(label = "Titles",  values = ~Titles_Won)
  )
) |> 
layout(autosize = FALSE, width = 750, height = 400,
       margin = list(
        l = 150,  # Left margin
        r = 50,   # Right margin (optional)
        b = 80,   # Bottom margin (optional)
        t = 120   # Top margin (optional)
      ),
      title = "EuroLeague Outcomes")



```

```{r}
#| echo: false
#| include: false

library(htmlwidgets)
library(webshot2)


p<- euro_plot_data |> 
plot_ly(
  type = 'parcoords',
  line = list(color = ~country_id,
              colorscale = 'Rainbow'),
  dimensions = list(
    list(label = "Team", values = ~team_id,
         tickvals = unique(euro_plot_data$team_id),
         ticktext = unique(euro_plot_data$Team)),
    list(label = "Country", values = ~country_id,
         tickvals = unique(euro_plot_data$country_id),
         ticktext = unique(euro_plot_data$Country)),
    list(label = "Capacity",  values = ~capacity),
    list(label = "2024-2025 Rank", values = ~lastrank,
         range = c(max(euro_plot_data$lastrank, na.rm = TRUE),
                   min(euro_plot_data$lastrank, na.rm = TRUE))),
    list(label = "FinalFour", values = ~FinalFour_Appearances),
    list(label = "Titles",  values = ~Titles_Won)
  )
) |> 
layout(autosize = FALSE, width = 750, height = 400,
       margin = list(
        l = 150,  # Left margin
        r = 50,   # Right margin (optional)
        b = 80,   # Bottom margin (optional)
        t = 120   # Top margin (optional)
      ),
      title = "EuroLeague Outcomes") 

saveWidget(p, "euroleague_thumbnail.html")
webshot("euroleague_thumbnail.html", file = "euroleague.png", vwidth = 750, vheight = 400)

```



```{r}
praise()
```

