---
title: "Henley Passport Index Data"
author: "Jo Hardin"
date: "09/09/2025"
format: html
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(jsonlite)
library(plotly)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```


## The Data

This week we are exploring data from the [Henley Passport Index API](api.henleypassportindex.com/api/v3/countries). 
The [Henley Passport Index](https://www.henleyglobal.com/passport-index/about) is produced by Henley & Partners and captures 
the number of countries to which travelers in possession of each passport in the world may enter visa free. 

```{r}
country_lists <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-09/country_lists.csv')
rank_by_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-09/rank_by_year.csv')

```




```{r}
travel_perms <- country_lists |> 
  pivot_longer(
    cols = c(visa_required, visa_online, visa_on_arrival, 
             visa_free_access, electronic_travel_authorisation),
    names_to = "visa_type",
    values_to = "json_str"
 ) |> 
  mutate(
    json_parsed = map(json_str, ~ fromJSON(.)[[1]])
) |> 
  unnest(json_parsed, names_sep = "_") |> 
  select(-json_str) |> 
  rename(from = code, from_name = country,
         to = json_parsed_code, to_name = json_parsed_name) |> 
  mutate(from = ifelse(from_name == "Namibia", "NA", from)) #|> 
 #mutate(visa_type = case_when(
 #  visa_type == "visa_required" ~ "visa_required",
 #  visa_type == "visa_online" ~ "visa_required",
 #  visa_type == "visa_on_arrival" ~ "visa_required",
 #  visa_type == "visa_free_access" ~ "easy_access",
 #  visa_type == "electronic_travel_authorisation" ~ "easy_access"
 #)) |> 
 #  distinct()

travel_perms
```

Adding the home countries so that they show up on the map.

```{r}
home_countries <- travel_perms |> 
  select(from, from_name) |> 
  distinct() |> 
  mutate(visa_type = "domestic",
         to = from,
         to_name = from_name)

travel_perms <- travel_perms |> rbind(home_countries)
```







## Plotting

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

world_data <- world |> 
  left_join(travel_perms, by = c("iso_a2_eh" = "to"))

world_data |> 
  filter(from == "MX") |> 
ggplot() +
  geom_sf(aes(fill = visa_type)) +
  theme_minimal() + 
  labs(title = "Visa connections for Mexicans")
```


```{r}
plot_ly() |> 
  add_sf(
    data = world_data |> filter(from == "MX"),
    split = ~iso_a2_eh,
    color = ~visa_type,
    text = ~paste0("Country: ", to_name, "<br>Visa: ", visa_type)#,
    #hoveron = "fills"
  ) |> 
  layout(title = "Visa connections for Mexicans",
         showlegend = FALSE)
```





```{r}
#| echo: false
#| include: false

png(filename = "passports.png", width = 550, height = 450)

world_data |> 
  filter(from == "MX") |> 
ggplot() +
  geom_sf(aes(fill = visa_type)) +
  theme_minimal() + 
  labs(title = "Visa connections for Mexicans")


dev.off()
```


```{r}
praise()
```