---
title: "Henley Passport Index Data"
author: "Jo Hardin"
date: "09/09/2025"
format: html
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(jsonlite)
library(plotly)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```


## The Data

This week we are exploring data from the [Henley Passport Index API](api.henleypassportindex.com/api/v3/countries). 
The [Henley Passport Index](https://www.henleyglobal.com/passport-index/about) is produced by Henley & Partners and captures 
the number of countries to which travelers in possession of each passport in the world may enter visa free. 

```{r}
country_lists <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-09/country_lists.csv')
rank_by_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-09/rank_by_year.csv')

```




```{r}
travel_perms <- country_lists |> 
  janitor::clean_names() %>%
  setNames(gsub("_", " ", names(.))) |> 
  pivot_longer(
    cols = c(`visa required`, `visa online`, `visa on arrival`, 
             `visa free access`, `electronic travel authorisation`),
    names_to = "visa_type",
    values_to = "json_str"
 ) |> 
  mutate(
    json_parsed = map(json_str, ~ fromJSON(.)[[1]])
) |> 
  unnest(json_parsed, names_sep = "_") |> 
  select(-json_str) |> 
  rename(from = code, from_name = country,
         to = json_parsed_code, to_name = json_parsed_name) |> 
  mutate(from = ifelse(from_name == "Namibia", "NA", from)) 
 #mutate(visa_type = case_when(
 #  visa_type == "visa_required" ~ "visa_required",
 #  visa_type == "visa_online" ~ "visa_required",
 #  visa_type == "visa_on_arrival" ~ "visa_required",
 #  visa_type == "visa_free_access" ~ "easy_access",
 #  visa_type == "electronic_travel_authorisation" ~ "easy_access"
 #)) |> 
 #  distinct()

travel_perms
```

Adding the home countries so that they show up on the map.

```{r}
home_countries <- travel_perms |> 
  select(from, from_name) |> 
  distinct() |> 
  mutate(visa_type = "domestic",
         to = from,
         to_name = from_name)

travel_perms <- travel_perms |> rbind(home_countries)
```







## Plotting

Would be great to have the plot be interactive so that the type of visa needed for travel is based on any country which is clicked on. Instead, we've created static map with information for Mexican travelers.

The interactive version of the plot (for any country's travel restrictions) is too big for **plotly**. It could be done using a **Shiny** app.

```{r}
#| fig.cap: "Each country is filled based on the type of visa needed for Mexican travelers. Note that Mexico is colored differently because domestic travel does not require a visa."
#| fig.alt: "World map with countries colored according to the type of visa needed for Mexican travelers."

world <- ne_countries(scale = "medium", returnclass = "sf")

world_data <- world |> 
  left_join(travel_perms, by = c("iso_a2_eh" = "to"))

world_data |> 
  filter(from == "MX") |> 
ggplot() +
  geom_sf(aes(fill = visa_type)) +
  theme_minimal() + 
  labs(title = "Visa connections for Mexicans",
       fill = "visa type")
```


```{r}
#| fig.cap: "Each country is filled based on the type of visa needed for Mexican travelers. Note that Mexico is colored differently because domestic travel does not require a visa."
#| fig.alt: "World map with countries colored according to the type of visa needed for Mexican travelers."

plot_ly() |> 
  add_sf(
    data = world_data |> filter(from == "MX"),
    split = ~iso_a2_eh,
    color = ~visa_type,
    text = ~paste0("Country: ", to_name, "<br>Visa: ", visa_type)#,
    #hoveron = "fills"
  ) |> 
  layout(title = "Visa connections for Mexicans",
         showlegend = FALSE)
```





```{r}
#| echo: false
#| include: false

png(filename = "passports.png", width = 550, height = 450)

world_data |> 
  filter(from == "MX") |> 
ggplot() +
  geom_sf(aes(fill = visa_type)) +
  theme_minimal() + 
  labs(title = "Visa connections for Mexicans",
       fill = "visa type")


dev.off()
```


```{r}
praise()
```