---
title: "Shakespeare Dialogue"
author: "Jo Hardin"
date: "09/17/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(scales)
library(tidytext)
library(devtools)
library(ggwordcloud)
library(png)
library(svglite)
```


## Data


This week we're exploring dialogue in Shakespeare plays. The dataset this week comes from [shakespeare.mit.edu](https://shakespeare.mit.edu/) (via [github.com/nrennie/shakespeare](https://github.com/nrennie/shakespeare)) which is the Web's first edition of the Complete Works of William Shakespeare. The site has offered Shakespeare's plays and poetry to the internet community since 1993.


```{r}
hamlet <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-17/hamlet.csv')
macbeth <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-17/macbeth.csv')
romeo_juliet <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-17/romeo_juliet.csv')
```


## Romeo & Juliet

After removing stop words, we create wordclouds describing the most common works for both Romeo and Juliet in *Romeo & Juliet*.  Tha anlysis is taken from <a href = "https://github.com/deepdk/TidyTuesday2024/tree/main/2024/week_38" target = "_blank">@deepdk</a>.


```{r}
romeo_juliet <- romeo_juliet |> 
  filter(character %in% c("Romeo", "Juliet"))
```


```{r}
# Create a custom list of words to exclude
custom_stop_words <- data.frame(word = "thou", "thy", "thee", "thine", "art", "hast", "dost", "ere", "o","hath")
```

```{r}
word_counts <- romeo_juliet |>
  unnest_tokens(word, dialogue) |>
  anti_join(stop_words) |>  # Remove common stop words
  filter(!str_detect(word, "^[0-9]+$")) |>  # Remove numbers
  anti_join(custom_stop_words) |>  # Remove custom words
  mutate(word = stringr::str_replace(word, "'s", "")) |> 
  count(character, word, sort = TRUE)

word_counts
```

```{r}
juliet <- word_counts |> 
  filter(character == "Juliet")

romeo <- word_counts |> 
  filter(character == "Romeo")
```

We wanted to use different fonts, so we load in the `MedievalSharp` font from Google.

```{r}
sysfonts::font_add_google("MedievalSharp", "MedievalSharp")
showtext::showtext_auto()

my_font <- "MedievalSharp"
```


:::: {.columns}

::: {.column width="50%"}
```{r}
#| fig-cap: "The most common words said by Juliet in Shakespeare's *Romeo & Juliet*"
juliet |> 
  #filter(n > 1) |> 
  ggplot(aes(label = word, size = n, color = n)) + 
  #ggwordcloud::geom_text_wordcloud(shape = "cardioid")
  ggwordcloud::geom_text_wordcloud_area(
    mask = png::readPNG("FlipAlphaShakespeare.png"),
    rm_outside = TRUE,
    family = my_font
  ) +
  scale_size_area(max_size = 20) +
  theme_minimal() +
  scale_color_gradient(low = "#03c6fc", high = "#5203fc") + 
  labs(title = "Shakespeare in Words") + 
  theme(
    plot.title = ggtext::element_textbox_simple(
      family = my_font),
    plot.caption = ggtext::element_textbox_simple(
      family = my_font)  )
```
:::

::: {.column width="50%"}
```{r}
#| fig-cap: "The most common words said by Romeo in Shakespeare's *Romeo & Juliet*"
romeo |> 
  #filter(n > 1) |> 
  ggplot(aes(label = word, size = n, color = n)) + 
  #ggwordcloud::geom_text_wordcloud(shape = "cardioid")
  ggwordcloud::geom_text_wordcloud_area(
    mask = png::readPNG("AlphaShakespeare.png"),
    rm_outside = TRUE,
    family = my_font
  ) +
  scale_size_area(max_size = 20) +
  theme_minimal() +
  labs(title = "Shakespeare in Words") + 
  theme(
    plot.title = ggtext::element_textbox_simple(
      family = my_font),
    plot.caption = ggtext::element_textbox_simple(
      family = my_font)  ) +
  scale_color_gradient(low = "#a5fc03", high = "#034efc")
```
:::
::::



```{r}
praise()
```



