---
title: "CIA Factbook"
author: "Jo Hardin"
date: "10/22/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(maps)
library(praise)
```


## The Data

This week we're exploring the [CIA World Factbook](https://www.cia.gov/the-world-factbook/)! 
The dataset this week comes from the [CIA Factbook, Country Comparisons, 2014](https://www.cia.gov/the-world-factbook/references/guide-to-country-comparisons),
via the [{openintro}](https://openintrostat.github.io/openintro/) R package,
via the [{usdatasets}](https://cran.r-project.org/package=usdatasets) R package, 
via [this post on LinkedIn](https://www.linkedin.com/posts/andrescaceresrossi_rstats-rstudio-opensource-activity-7249513444830318592-r395).

> The *World Factbook* provides basic intelligence on the history, people, government, 
> economy, energy, geography, environment, communications, transportation, military, 
> terrorism, and transnational issues for 265 world entities.

```{r}
cia <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-22/cia_factbook.csv') |> 
  mutate(iso_a3 = countrycode::countrycode(
      country,
      origin = "country.name",
      destination = "iso3c"
    )) |> 
  drop_na(iso_a3)
```


## World Map

```{r}
world_data <- map_data("world") |>   # ggplot2
  mutate(iso_a3 = countrycode::countrycode(
      region,
      origin = "country.name",
      destination = "iso3c"
    )) |> 
  drop_na(iso_a3)

data4map <- world_data |> 
  left_join(cia, by = "iso_a3")
```


```{r}
#| fig-cap: world map colored by the death rate (number of deaths per 1,000 people) in 2014.
#| fig-alt: world map where each country is colored by the number of deaths per 1000 people in 2014.
data4map |> 
  ggplot(aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = death_rate), color = "black", size = 0.15) + 
  theme_void() + 
  scale_fill_gradient(trans = "log")
  #scale_fill_gradient(low = "blue", high = "purple")
```

## Shiny App

I also created a Shiny App which allows the user to change the variable that colors each country.  However, I don't have an obvious place to host the Shiny App.  The source code for the Shiny App is in `app.R` which is located at https://github.com/hardin47/TidyTuesday/tree/gh-pages/2024-10-22/CIAFactbook .



```{r}
praise()
```

