---
title: "Hermann's Tortoises"
author: "Jo Hardin"
date: March 3, 2026
format: 
  html:
    embed-resources: true
    code-fold: true
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## The Data

### Golem Grad Tortoise Data

The datasets this week come from the paper "[Sex Ratio Bias Triggers Demographic Suicide in a Dense Tortoise Population](https://onlinelibrary.wiley.com/doi/10.1111/ele.70296)".
This topic seemed so interesting that even media like the New York Times picked it up:
"[Constant Sexual Aggression Drives Female Tortoises to Walk Off Cliffs](https://www.nytimes.com/2026/02/14/science/tortoises-island-sex-cliff.html)".

> In an exceptionally dense island population of Hermann's tortoises 
> in Lake Prespa in North Macedonia, sexually coercive males 
> dramatically overnumber females, inflict severe copulatory injuries 
> and put them at risk of fatal falls from the island plateau's sheer 
> rock faces. Harassed females are emaciated, reproduce less frequently,
> produce smaller clutches and have lower annual survival rates compared 
> to females from a neighbouring mainland population. Sixteen years of 
> capture-recapture data reveal an ongoing extinction event and predict
> that the last island female will die in 2083. 

- Do recaptures happen more often in spring or summer? 
- Does it seem easier to recapture male or female tortoises?
- What are the differences among tortoises from the mainland vs the ones from the island in terms of body mass or carapace length?

Thank you to [Novica Nakov, Free Software Macedonia](https://github.com/novica) for curating this week's dataset.



```{r}
clutch_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-03-03/clutch_size_cleaned.csv')
tortoise <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-03-03/tortoise_body_condition_cleaned.csv')
```



```{r}
clutch_size |> 
  ggplot() +
  geom_point(aes(x = body_mass_grams, y = eggs, color = age)) + 
  facet_wrap((~locality))
```

```{r}
tortoise |> 
  group_by(individual) |> 
  summarize(count = n()) |> 
  arrange(desc(count))
```

```{r}
tortoise |> 
  filter(individual == 644) |> 
  ggplot() + 
  geom_point(aes(x = year, y = straight_carapace_length_mm)) 
```

```{r}
tortoise |> 
  group_by(individual) |> 
  summarize(count = n(), locality = first(locality), sex = first(sex))
```



```{r}
tortoise |> 
  group_by(year, locality) |> 
  mutate(sex_ratio = mean(sex == "f")) |> 
  ggplot() + 
  geom_point(aes(x = sex_ratio, y = body_mass_grams, color = sex), 
             alpha = 0.5, size = 0.5) + 
  facet_wrap(~locality, scales = "free_x")
```



```{r}
tortoise |> 
  ggplot() + 
  geom_point(aes(x = body_mass_grams, y = body_condition_index, color = sex))
```


```{r}
tortoise |> 
  mutate(new_condition = body_mass_grams / straight_carapace_length_mm) |> 
  select(body_mass_grams, straight_carapace_length_mm, body_condition_index, new_condition)
```

We expect the body condition of females to be higher.  Turns out that the female tortoises are only perceptibly larger in the mainland, Konjsko.


```{r}
#| fig.cap: "Comparing the body condition of Hermann's tortoises at / near Golem Grad, an island near Macedonian. We expect female tortoises to have a higher body condition, but that is seen only on Konjsko mainland."
#| fig.alt: "Violin plot of the body condition of tortoises broken down by sex and location. Only Konjsko on the mainland has the body condition of female tortoises to be substantially higher than male tortoises. On the beach and plateau, the body conditions across genders is quite similar."

library(ggpattern) 

tortoise |> 
  mutate(locality = forcats::fct_relevel(locality, c("Konjsko", "Beach", "Plateau")),
         sex = ifelse(sex == "f", "F", "M")) |> 
  group_by(sex, locality) |> 
  mutate(ave_condition = mean(body_condition_index)) |> 
  ggplot() + 
  geom_violin_pattern(aes(y = body_condition_index, x = sex,
                  fill = locality, pattern = sex)) +
  geom_boxplot(aes(y = body_condition_index, x = sex),
               whisker.linewidth = 0, median.linewidth = 0,
               width = .2, outliers = FALSE)  +
  geom_point(aes(y = ave_condition, x = sex), 
             shape = "+", size = 5) +
  scale_pattern_manual(values = c("stripe", "none")) + 
  scale_fill_manual(values = c("#2F4F2F", "#6B8E23", "#A3B86C")) + 
  labs(pattern = "Sex", fill = "Locality", x = "", y = "Body condition") + 
  guides(fill = guide_legend(override.aes = list(pattern = "none")))  + 
  facet_grid(~ locality, switch = "x") + 
  theme_minimal()
```


```{r}
#| echo: false
#| include: false

p <- tortoise |> 
  mutate(locality = forcats::fct_relevel(locality, c("Konjsko", "Beach", "Plateau")),
         sex = ifelse(sex == "f", "F", "M")) |> 
  group_by(sex, locality) |> 
  mutate(ave_condition = mean(body_condition_index)) |> 
  ggplot() + 
  geom_violin_pattern(aes(y = body_condition_index, x = sex,
                  fill = locality, pattern = sex)) +
  geom_boxplot(aes(y = body_condition_index, x = sex),
               whisker.linewidth = 0, median.linewidth = 0,
               width = .2, outliers = FALSE)  +
  geom_point(aes(y = ave_condition, x = sex), 
             shape = "+", size = 5) +
  scale_pattern_manual(values = c("stripe", "none")) + 
  scale_fill_manual(values = c("#2F4F2F", "#6B8E23", "#A3B86C")) + 
  labs(pattern = "Sex", fill = "Locality", x = "", y = "Body condition") + 
  guides(fill = guide_legend(override.aes = list(pattern = "none")))  + 
  facet_grid(~ locality, switch = "x") + 
  theme_minimal()


png(filename = "tortoise.png", width = 550, height = 450)

p

dev.off()

```


```{r}
praise()
```
