---
title: "International Mathematics Olympiad"
author: "Jo Hardin"
date: "09/24/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(maps)
library(gganimate)
```



## The Data

The data for this week comes from International Mathematical Olympiad (IMO). Thank you to [Havisha Khurana](https://github.com/havishak) for curating this week's dataset.

> The International Mathematical Olympiad (IMO) is the World Championship Mathematics Competition for High School students and is held annually in a different country. The first IMO was held in 1959 in Romania, with 7 countries participating. It has gradually expanded to over 100 countries from 5 continents. The competition consists of 6 problems and is held over two consecutive days with 3 problems each.


```{r}
country_results_df <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-24/country_results_df.csv')
individual_results_df <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-24/individual_results_df.csv')
timeline_df <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-24/timeline_df.csv')
```


In 2024, there were 609 people competing from 109 different countries.

```{r}
indiv_country <- individual_results_df |> 
  mutate(country2 = ifelse(str_detect(country, "C[0-9]"), "Russia", country)) 

indiv_country |> 
  filter(year == 2024) |> 
  summarise(distinct_countries = n_distinct(country2), n_countries = n())
```


```{r}
indiv_country <- indiv_country |> 
  group_by(country2, year) |> 
  arrange(individual_rank) |> 
  slice_head(n = 1) |> 
  arrange(desc(year)) |> 
  select(year, country2, individual_rank, award) |> 
  mutate(country3 = case_when(
    country2 == "United States of America" ~ "USA",
    country2 == "People's Republic of China" ~ "China",
    country2 == "Türkiye" ~ "Turkey",
    country2 == "Islamic Republic of Iran" ~ "Iran",
    country2 == "Republic of Korea" ~ "South Korea",
    country2 == "Islamic Republic of Iran" ~ "Iran",
    country2 == "Turkish Republic of Northern Cyprus" ~ "Cyprus",
    country2 == "Democratic People's Republic of Korea" ~ "North Korea",
    TRUE ~ country2
  )) |> 
  mutate(award = as.factor(case_when(
         str_detect(award, "Gold") ~ "Gold",
         str_detect(award, "Silver") ~ "Silver",
         str_detect(award, "Bronze") ~ "Bronze",
         TRUE ~ NA))) |> 
  mutate(award = fct_relevel(award, c("Gold", "Silver", "Bronze")))

indiv_country
```


## Making a map

```{r}
world_data <- map_data("world")

world_data |> 
  summarize(n_distinct(region))

world_data |> 
  inner_join(indiv_country, by = c("region" = "country3")) |> 
  summarize(n_distinct(region))

plot_imo <- world_data |> 
  left_join(indiv_country, by = c("region" = "country3")) 
```



```{r}
#| fig-cap: "For each year, the rank of the top performing student in each country on the International Mathematical Olympiad."
#| fig-alt: "A map of the world.  Each country is colored based on the rank of the top performing student on the International Mathematical Olympiad. The country is outlined in either gold, silver, or bronze depending if the top performing student won a gold, silver, or bronze medal.  The plot is animated to change over years."
library(gganimate)
my_breaks <- c(1, 10, 50, 200, 600)

plot_imo |> 
  mutate(year = as.factor(year)) |> 
  #filter(year == 1979 | !is.na(individual_rank)) |> 
  filter(!is.na(individual_rank)) |> 
ggplot(aes(x = long, y = lat, group = group)) + 
  geom_polygon(data = world_data, 
               aes(x = long, y = lat, group = group),
               color = "black", fill = "white") + 
  geom_polygon(aes(fill = individual_rank, color = award)) + 
  scale_fill_gradient(name = "best individual rank", 
                      trans = "log",
                      low = "#71797E",
                      high = "#E6E6E6",
                      breaks = my_breaks, labels = my_breaks) + 
  scale_color_manual(values = c("#FFD700", "#C0C0C0", "#CE8946"),
                     na.value="black",
                     breaks = c("Gold", "Silver", "Bronze")) +
  gganimate::transition_states(year, 
                               state_length = 5,
                               transition_length = 10) +
  labs(title = 'Math Olympiad top scorer in {closest_state}')

```



```{r}
praise()
```

