---
title: "ISO Country Codes"
author: "Jo Hardin"
date: "11/12/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(plotly)
library(praise)
```


## The Data

We've referenced countries and country codes in many past datasets, but we've never looked closely at the ISO 3166 standard that defines these codes.

Wikipedia says:

> ISO 3166 is a standard published by the International Organization for 
> Standardization (ISO) that defines codes for the names of countries, dependent 
> territories, special areas of geographical interest, and their principal 
> subdivisions (e.g., provinces or states). The official name of the standard 
> is Codes for the representation of names of countries and their subdivisions.

The dataset this week comes from the {[ISOcodes](https://cran.r-project.org/package=ISOcodes)} R package. It consists of three tables:

- `countries`: Country codes from ISO 3166-1.
- `country_subdivisions`: Country subdivision code from ISO 3166-2.
- `former_countries`: Code for formerly used names of countries from ISO 3166-3.


```{r}
countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/countries.csv')
country_subdivisions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/country_subdivisions.csv')
former_countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/former_countries.csv')

```


```{r}
former <- former_countries |> 
  mutate(date = case_when(
    nchar(date_withdrawn) == 4 ~ ymd(paste0(date_withdrawn, "-01-01")),
    TRUE ~ ymd(date_withdrawn)
  )) |> 
  mutate(country = fct_reorder(alpha_3, date))
```



```{r}
country_subdivisions |> 
  group_by(type) |> 
  summarize(count_type = n()) |> 
  arrange(desc(count_type))
```



```{r}
#| fig-cap: Thirty-one countries have been withdrawn from the ISO 3166 standard in recent years.  Hover on the dot to see the name of the country and the year they were withdrawn.
#| fig-alt: A horizonal lollipop plot with date on the x axis and country on the y axis.  The graph is interactive so that when you hover on the lollipop, the country name and year they were withdrawn is observed.


p <- former |> 
  ggplot(aes(x = date, y = country)) + 
  geom_segment(aes(xend = min(date), yend = country)) + 
  geom_point(aes(
    text = paste("Country: ", name, "<br>Date: ", date_withdrawn,
                 "<br>Info: ", comment)), 
    size = 3) + 
  theme_minimal() +
  scale_y_discrete(limits=rev) + 
  labs(title = "Date of Country Withdrawn",
       y = "", x = "Date")

ggplotly(p, tooltip = "text")

```



