---
title: "Market Indicators"
author: "Jo Hardin"
date: August 6, 2026
format: 
  html:
    embed-resources: true
    code-fold: false
execute:
  warning: false
  message: false
---

Today we are measuring different market indicators using data from the <a href = "https://github.com/datasets" target = "_balnk">Curated open data on GitHub</a>.


```{r}
library(tidyverse)
library(praise)
```


## Data

We collect data over time on: gold prices, natural gas prices, oil prices, the yield rate of the US 10y bond, volatility index, and the S & P 500.

```{r}
#| code-fold: true
gold <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/gold-prices/refs/heads/main/data/monthly-processed.csv'
) |>
  rename(gold_price = Price)

oil <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/oil-prices/refs/heads/main/data/wti-monthly.csv'
) |>
  rename(oil_price = Price)

nat_gas <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/natural-gas/refs/heads/main/data/monthly-processed.csv'
) |>
  rename(gas_price = Price)

vix <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/finance-vix/refs/heads/main/data/vix-monthly.csv'
) |> # volatility index
  rename(vol_idx = Close)

bond <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/bond-yields-us-10y/refs/heads/main/data/monthly.csv'
) |> # bond yields us 10y
  rename(bond_rate = Rate)

temp <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/global-temp/refs/heads/main/data/monthly.csv'
) |>
  mutate(Date = lubridate::as_date(paste0(Year, "-01"))) |>
  rename(temp_diff = Mean) |>
  select(-Year, -Source)
# difference of that month from the average over all of the 20th century

sp <- readr::read_csv(
  'https://raw.githubusercontent.com/datasets/s-and-p-500/refs/heads/main/data/data.csv'
) |>
  select(Date, SP500)

full_data <- gold |>
  full_join(oil) |>
  full_join(nat_gas) |>
  full_join(vix) |>
  full_join(bond) |>
  full_join(temp) |>
  full_join(sp)
```


## Making a single plot

We start by making a single line plot over time, measuring oil prices.

```{r}

full_data |>
  select(Date, oil_price) |>
  drop_na() |>
  ggplot(aes(x = Date, y = oil_price)) +
  geom_line() +
  theme_bw()
```


## Write a function

The function takes a variable name as the argument and makes a line plot over time for that variable.  Note the `{{ }}` needed due to the non-standard evaluation.

```{r}
plot_ts <- function(variable, log = TRUE) {
  if (log) {
    full_data |>
      select(Date, {{ variable }}) |>
      drop_na() |>
      ggplot(aes(x = Date, y = {{ variable }})) +
      geom_line() +
      theme_bw() +
      scale_y_log10() +
      labs(x = "date")
  } else {
    full_data |>
      select(Date, {{ variable }}) |>
      drop_na() |>
      ggplot(aes(x = Date, y = {{ variable }})) +
      geom_line() +
      theme_bw() +
      labs(x = "date")
  }
}

plot_ts(SP500, log = TRUE)
plot_ts(SP500, log = FALSE)
```

## Mapping the function

Unforunately, the non-standard evaluation means that the function we wrote cannot be directly mapped.  `map()` will go through the variables vector one at a time (e.g., `variables[1]`, `variables[2]`, etc.), and the name of the variable then needs to be translated into something that can be used in the tidy pipeline.


```{r}
variables <- c("vol_idx", "gold_price", "oil_price")

map(
  variables,
  \(v, log = TRUE) plot_ts(!!rlang::sym(v), log = log),
  log = FALSE
)
```

## Plotting multiple variables


Instead of looking at a single variable at a time, we might be interested in looking at multiple variables on the same plot. It is a little bit tricky because the variables are meausured on quite different scales, but some are comparable.

```{r}
#| fig-cap: "Comparing the US 10y yield bond rate, the S&P 500, and global temperature deviations from the average temperature for all of the 20th century."
#| fig-alt: "A line plot with time from the 1950s to the 2020s on the x-axis. The y-axis measures three distinct variables: (1) the US 10y yield bond rate, (2) the S and P 500, and (3) the global temperature deviations from the average temperature for all of the 20th centure. The three time series are not particularly correlated. The bond rate peaks sharply in the 1980s and has started to rise again in the 2020. The S and P has risen a lot over the time period and is numerically much later than the other two time series. The temperature has risen slowly over the entire time period."

full_data |>
  select(Date, bond_rate, temp_diff, SP500) |>
  drop_na() |>
  pivot_longer(
    cols = -Date,
    names_to = "product",
    values_to = "value"
  ) |>
  ggplot(aes(x = Date, y = value, color = product)) +
  geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Dark2") +
  scale_y_log10() +
  labs(y = "", title = "Comparison over time, log10 scale")
```

## Using Shiny

We also played around with Shiny so the user could choose which variable to plot:

<iframe src="https://019fdd7a-0f99-982e-6893-85608fd3a282.share.connect.posit.cloud/" width="90%" height="800"   style="border: none;" data-external="1"></iframe>

```{r}
#| echo: false
#| include: false

p <- full_data |>
  select(Date, bond_rate, temp_diff, SP500) |>
  drop_na() |>
  pivot_longer(
    cols = -Date,
    names_to = "product",
    values_to = "value"
  ) |>
  ggplot(aes(x = Date, y = value, color = product)) +
  geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Dark2") +
  labs(y = "", title = "Comparison over time, log10 scale") +
  scale_y_log10()


png(filename = "marketindex.png", width = 650, height = 450)

p

dev.off()
```


```{r}
praise()
```
