---
title: "UK Meteorological & Climate Data"
author: "Jo Hardin"
date: October 21, 2025
format: 
  html:
    embed-resources: true
    code-fold: true
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(shiny)
```


## The Data


This week we're exploring [historic meteorological
data](https://www.metoffice.gov.uk/research/climate/maps-and-data/historic-station-data)
from the [UK Met Office](https://www.metoffice.gov.uk/).

The UK Met Office is the United Kingdom’s national weather and climate service,
providing forecasts, severe weather warnings, and climate science expertise. It
helps people, businesses, and governments make informed decisions to stay safe
and plan for the future. It was first established in 1854, making it one of the
oldest weather services in the world.

Data has been scraped straight from the Met Office website and cleaned in a
basic way. The few flags for "estimated" data and changing sunlight monitoring
techniques have been removed for simplicity, but are available from the raw
data. Also available is simple site metadata which includes the opening date and
[latitude and longitude](https://epsg.io/4326) of each station.

> Monthly Historical information for 37 UK Meteorological Stations. Most go back
to the early 1900s, but some go back as far as 1853. Station data files are
updated on a rolling monthly basis, around 10 days after the end of the month.
No allowances have been made for small site changes and developments in
instrumentation.


```{r}
historic_station_met <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-10-21/historic_station_met.csv')
station_meta <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-10-21/station_meta.csv')
```


## Spiral plot

For each month of each year, we can plot the amount of rain or sun or temperature. Notice that amount of sun hasn't changed much over the years, but rain and temperature have. That is, the darkest arc for high and low temperature in any piece of pie is toward the later years. Rain is sort of all over the place, it doesn't really show that rain is necessarily decreasing (or increasing).


### High temperature (in Aberporth)



```{r}
month_length <- c(31, 28, 31, 30, 31, 30,
                  31, 31, 30, 31, 30, 31)

month_breaks <- cumsum(month_length) - 30
base_grey <- "grey28"

year_annotations <- list(
  year = c(1960, 1980, 2000, 2020),
  x = rep(3, 4),
  y = ymd(paste(c(1960, 1980, 2000, 2020), "01", "01", sep = "-"))
)
```

```{r}
#| fig.alt: "High temperatures recorded for each month and each year since about 2000 in Aberporth. The values are presented in a spiral so that we can see how the temperature changes across year while keeping month constant. There are slighlty higher temperatures in recent years across each month."

historic_station_met |> 
  filter(station == "aberporth") |> 
  filter(year >= 2000) |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = tmax), linewidth = 2) +
  scale_color_gradient(low = "white", high = "darkgreen") +
  annotate("text", 
           label = paste0(year_annotations$year, "\u2192"), 
           x = year_annotations$x, 
           y = year_annotations$y, 
           family = "Arial",
           size = 2, hjust = c(1.1, 1.2, 1.3, 1.4)) +#, hjust = 0.15) +

  coord_polar() + 
  theme_void() +
  scale_x_continuous(minor_breaks = month_breaks, 
                     breaks = month_breaks[c(3, 6, 9, 12)],
                     labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  ) + 
  labs(title = "High temperature (in Aberporth)")
```

### Low temperature (in Aberporth)

```{r}
historic_station_met |> 
  filter(station == "aberporth") |> 
  filter(year >= 2000) |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = tmin), linewidth = 2) +
  scale_color_gradient(low = "white", high = "maroon") +
  annotate("text", 
           label = paste0(year_annotations$year, "\u2192"), 
           x = year_annotations$x, 
           y = year_annotations$y, 
           family = "Arial",
           size = 2, hjust = c(1.1, 1.2, 1.3, 1.4)) +#, hjust = 0.15) +

  coord_polar() + 
  theme_void() +
  scale_x_continuous(minor_breaks = month_breaks, 
                     breaks = month_breaks[c(3, 6, 9, 12)],
                     labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  ) + 
  labs(title = "Low temperature (in Aberporth)")
```



### Rain (in Aberporth)

```{r}
historic_station_met |> 
  filter(station == "aberporth") |> 
  filter(year >= 2000) |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = rain), linewidth = 2) +
  scale_color_gradient(low = "white", high = "darkblue") +
  annotate("text", 
           label = paste0(year_annotations$year, "\u2192"), 
           x = year_annotations$x, 
           y = year_annotations$y, 
           family = "Arial",
           size = 2, hjust = c(1.1, 1.2, 1.3, 1.4)) +#, hjust = 0.15) +

  coord_polar() + 
  theme_void() +
  scale_x_continuous(minor_breaks = month_breaks, 
                     breaks = month_breaks[c(3, 6, 9, 12)],
                     labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  ) + 
  labs(title = "Rain (in Aberporth)")
```



```{r}
historic_station_met |> 
  filter(station == "aberporth") |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = sun), linewidth = 2) +
  scale_color_gradient(low = "yellow", high = "red", na.value = "white") +
  annotate("text", 
           label = paste0(year_annotations$year, "\u2192"), 
           x = year_annotations$x, 
           y = year_annotations$y, 
           family = "Arial",
           size = 2, hjust = c(1.1, 1.2, 1.3, 1.4)) +#, hjust = 0.15) +
  coord_polar() + 
  theme_void() +
  scale_x_continuous(minor_breaks = month_breaks, 
                     breaks = month_breaks[c(3, 6, 9, 12)],
                     labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  ) + 
  labs(title = "Sun (in Aberporth)")
```

Sadly, plotly won't work with `coord_polar()`. So we can't create a pull down menu for the station. We could if we used shiny, but that isn't in the cards for today. 

```{r}
praise()
```


```{r}
#| echo: false
#| include: false

png(filename = "climate.png", width = 550, height = 450)

historic_station_met |> 
  filter(station == "aberporth") |> 
  filter(year >= 2000) |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = tmax), linewidth = 2) +
  scale_color_gradient(low = "white", high = "darkgreen") +
  annotate("text", 
           label = paste0(year_annotations$year, "\u2192"), 
           x = year_annotations$x, 
           y = year_annotations$y, 
           family = "Arial",
           size = 2, hjust = c(1.1, 1.2, 1.3, 1.4)) +#, hjust = 0.15) +

  coord_polar() + 
  theme_void() +
  scale_x_continuous(minor_breaks = month_breaks, 
                     breaks = month_breaks[c(3, 6, 9, 12)],
                     labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  ) + 
  labs(title = "High temperature (in Aberporth)")


dev.off()
```



```{r}
praise()
```

