---
title: "UK Meteorological & Climate Data"
author: "Jo Hardin"
format: 
  html:
    embed-resources: true
runtime: shiny
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(shiny)
```


## The Data


This week we're exploring [historic meteorological
data](https://www.metoffice.gov.uk/research/climate/maps-and-data/historic-station-data)
from the [UK Met Office](https://www.metoffice.gov.uk/).

The UK Met Office is the United Kingdom’s national weather and climate service,
providing forecasts, severe weather warnings, and climate science expertise. It
helps people, businesses, and governments make informed decisions to stay safe
and plan for the future. It was first established in 1854, making it one of the
oldest weather services in the world.

Data has been scraped straight from the Met Office website and cleaned in a
basic way. The few flags for "estimated" data and changing sunlight monitoring
techniques have been removed for simplicity, but are available from the raw
data. Also available is simple site metadata which includes the opening date and
[latitude and longitude](https://epsg.io/4326) of each station.

> Monthly Historical information for 37 UK Meteorological Stations. Most go back
to the early 1900s, but some go back as far as 1853. Station data files are
updated on a rolling monthly basis, around 10 days after the end of the month.
No allowances have been made for small site changes and developments in
instrumentation.


```{r}
historic_station_met <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-10-21/historic_station_met.csv')
station_meta <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-10-21/station_meta.csv')
```





```{r}
historic_station_met |> 
  filter(station == "aberporth") |> 
  filter(year >= 2000) |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = rain), linewidth = 2) +
  #scale_color_gradient(low = "white", high = "darkblue") +
  scale_color_gradient2("BlPu") + 
  coord_polar() + 
  theme_void()
```


```{r}
month_length <- c(31, 28, 31, 30, 31, 30,
                  31, 31, 30, 31, 30, 31)

month_breaks <- cumsum(month_length) - 30
base_grey <- "grey28"

year_annotations <- list(
  year = c(1960, 1980, 2000, 2020),
  x = rep(3, 4),
  y = ymd(paste(c(1960, 1980, 2000, 2020), "01", "01", sep = "-"))
)
```

```{r}
historic_station_met |> 
  filter(station == "aberporth") |> 
  #filter(year == 2000) |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         # Handle December specially - extend to end of year instead of next year
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) |> 
  ggplot() +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = sun), linewidth = 2) +
  scale_color_gradient(low = "yellow", high = "red", na.value = "white") +
  annotate("text", 
           label = paste0(year_annotations$year, "\u2192"), 
           x = year_annotations$x, 
           y = year_annotations$y, 
           family = "Arial",
           size = 2, hjust = c(1.1, 1.2, 1.3, 1.4)) +#, hjust = 0.15) +
  coord_polar() + 
  theme_void() +
  scale_x_continuous(minor_breaks = month_breaks, 
                     breaks = month_breaks[c(3, 6, 9, 12)],
                     labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  )
```


```{r}
#| eval: false
data <- historic_station_met
categories <- levels(as.factor(historic_station_met$station))

# Create one ggplotly trace per category
plots <- lapply(categories, function(cat) {
  df_cat <- historic_station_met |> 
  #filter(station == as.character(cat)) |> 
  filter(station == "aberporth") |> 
  mutate(date = ymd(paste(year, month, "01", sep = "-")),
         day_of_year = yday(date),
         ymd_month = if_else(month == 12, 
                             ymd(paste(year, "12", "31", sep = "-")),
                             date %m+% months(1)),
         day_month = yday(ymd_month)) 

  p_cat <- ggplot(df_cat) +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = as.numeric(date), yend = as.numeric(ymd_month), color = sun), linewidth = 2) +
  scale_color_gradient(low = "yellow", high = "red", na.value = "white") +
  coord_polar() + 
  theme_void() +
  #scale_x_continuous(minor_breaks = month_breaks, 
  #                   breaks = month_breaks[c(3, 6, 9, 12)],
  #                   labels = c("March", "June", "Sep.", "Dec.")) +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  )
   ggplotly(p_cat)
})
```

```{r}
#| eval: false

# Extract traces (plotly objects can be lists of traces)
all_traces <- sapply(plots, function(x) x$x$data)
all_traces <- unlist(all_traces, recursive = FALSE)

# Create initial figure with all traces
fig <- plotly::plot_ly()
for (tr in all_traces) fig$x$data <- append(fig$x$data, list(tr))

# Define visibility (only the first category visible)
visible_list <- rep(FALSE, length(all_traces))
visible_list[1:length(plots[[1]]$x$data)] <- TRUE

fig <- fig |>
  layout(
    updatemenus = list(list(
      buttons = lapply(seq_along(categories), function(i) {
        vis <- rep(FALSE, length(all_traces))
        trace_counts <- vapply(plots, function(x) length(x$x$data), numeric(1))
        start_idx <- sum(trace_counts[seq_len(i - 1)]) + 1
        end_idx <- sum(trace_counts[seq_len(i)])
        idx <- seq(start_idx, end_idx)
        vis[idx] <- TRUE
        list(
          method = "update",
          args = list(list(visible = vis)),
          label = as.character(categories[i])
        )
      }),
      direction = "down",
      x = 0.1, y = 1.15
    ))
  )


fig
```




```{r}
#| eval: false

library(ggplot2)
library(plotly)
library(dplyr)
library(lubridate)

df_cat <- historic_station_met |> 
  filter(station == "aberporth") |> 
  mutate(
    date = ymd(paste(year, month, "01", sep = "-")),
    day_of_year = yday(date),
    ymd_month = if_else(
      month == 12, 
      ymd(paste(year, "12", "31", sep = "-")),
      date %m+% months(1)
    ),
    day_month = yday(ymd_month)
  )

p_cat <- ggplot(df_cat) +
  geom_segment(aes(x = day_of_year, xend = day_month, 
                   y = date, yend = ymd_month, color = sun), linewidth = 2) +
  scale_color_gradient(low = "yellow", high = "red", na.value = "white") +
  coord_polar() + 
  theme_void() +
  theme(
    plot.background = element_rect(color = NA, fill = "white"),
    panel.grid.major.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    panel.grid.minor.x = element_line(color = "grey", size = .5, linetype = "dotted"),
    axis.text.x = element_text(color = base_grey, size = 10)
  )

ggplotly(p_cat)

```


```{r}
selectInput(
inputId = "station",
label = "Choose station:",
choices = unique(historic_station_met$station)
)


renderPlot({
df_cat <- historic_station_met %>%
filter(station == input$station) %>%
mutate(
date = ymd(paste(year, month, "01", sep = "-")),
day_of_year = yday(date),
ymd_month = if_else(
month == 12,
ymd(paste(year, "12", "31", sep = "-")),
date %m+% months(1)
),
day_month = yday(ymd_month)
)

ggplot(df_cat) +
geom_segment(
aes(
x = day_of_year, xend = day_month,
y = as.numeric(date), yend = as.numeric(ymd_month),
color = sun
),
linewidth = 2
) +
scale_color_gradient(low = "yellow", high = "red", na.value = "white") +
coord_polar() +
theme_void() +
theme(
plot.background = element_rect(color = NA, fill = "white"),
panel.grid.major.x = element_line(color = "grey", size = 0.5, linetype = "dotted"),
panel.grid.minor.x = element_line(color = "grey", size = 0.5, linetype = "dotted"),
axis.text.x = element_text(color = base_grey, size = 10)
)
})



```



```{r}
praise()
```




