---
title: "National Park Species"
author: "Jo Hardin"
date: "10/08/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
library(waffle)
```


## The Data

This week we're exploring species at the most visited National Parks in the USA! NPSpecies contains species listed by National Parks maintained by National Parks Service (NPS). Given the size of the dataset, we're focusing on the 15 most visited parks. The data comes from https://irma.nps.gov/NPSpecies/Search/SpeciesList. 

> The information in NPSpecies is available to the public. The exceptions to this are records for some sensitive, threatened, or endangered species, where widespread distribution of information could potentially put a species at risk.   
>
> An essential component of NPSpecies is evidence; that is, observations, vouchers, or reports that document the presence of a species in a park. Ideally, every species in a park that is designated as “present in park” will have at least one form of credible evidence substantiating the designation

If you are looking for more detailed information on the dataset, here is the glossary for column names, field options, and tag meanings: 
https://irma.nps.gov/content/npspecies/Help/docs/NPSpecies_User_Guide.pdf


To properly cite NPSpecies use the following: 
**NPSpecies - The National Park Service biodiversity database.**
**https://irma.nps.gov/npspecies/. Accessed date/time. **

*This data was accessed on September 2nd, 2024.*

If you are interested in additional data, the curated dataset for all national parks is available at [https://github.com/frankiethull/NPSpecies](https://github.com/frankiethull/NPSpecies).

Thank you to [f. hull](https://github.com/frankiethull) for curating this week's dataset.


```{r}
species <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-08/most_visited_nps_species_data.csv')
```

```{r}
species |> 
  distinct(ParkCode)
```



```{r}
species |> 
  group_by(ParkCode, Occurrence) |> 
  summarize(biodiv = n()) |> 
  ggplot(aes(x = ParkCode, y = biodiv, color = Occurrence)) + 
  geom_point() + 
  scale_y_log10() +
  coord_flip()

species |> 
  group_by(ParkCode, Occurrence) |> 
  summarize(biodiv = n()) |> 
  ggplot(aes(x = ParkCode, y = biodiv, color = Occurrence)) + 
  geom_point() + 
  coord_flip()
```
We filter so that the graph is easier to see.  We consider only Joshua Tree (JOTR),
Olympic (OLYM), and Yosemite (YOSE).

```{r}
#| fig-cap: Each box represents 10 species in a particular category.  All three parks have a biodiversity dominated by Vascular Plants.
#| fig-alt: A waffle plot with national park on the x-axis (Joshua Tree, Olympic, or Yosemite) and number of species on the y-axis. The waffle squares are colored by the category type (vascular plant, reptile, mammal, insect, fungi, fish, bird, etc.). All three parks are dominated by multiple species of vasular plants. 

color_palette <- c(
  "Mammal" = "#8B4513",        # Dark brown
  "Bird" = "#FFA500",          # Bright orange
  "Reptile" = "#6B8E23",       # Olive green
  "Amphibian" = "#32CD32",     # Lime green
  "Fish" = "#20B2AA",          # Teal
  "Vascular Plant" = "#228B22",# Forest green
  "Spider/Scorpion" = "#800080",# Dark purple
  "Insect" = "#FFD700",        # Golden yellow
  "Other Non-vertebrates" = "#A9A9A9", # Gray
  "Fungi" = "#8B0000",         # Dark red
  "Bacteria" = "#000080"       # Navy blue
)

species |> 
  #filter(Nativeness %in% c("Native", "Non-native")) |> 
  count(ParkCode, ParkName, CategoryName) |> #, Nativeness) |> 
  filter(n >= 10) |> 
  filter(ParkCode %in% c("JOTR", "OLYM", "YOSE")) |> 
  mutate(n = round(n /10)) |> 
  ggplot(aes(fill = CategoryName, values = n)) +
  geom_waffle(color = "white", size = 0.25, n_rows = 10, flip = TRUE) + 
  facet_grid( ~ ParkName) + 
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 100, 
                     expand = c(0,0)) +
  coord_equal() +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE)) + 
  scale_fill_manual(values = color_palette)
```

```{r}
species |> 
  distinct(CategoryName)
```





```{r}
praise()
```

