---
title: "UN Votes"
author: "Jo Hardin"
date: "3/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width=10)
library(tidyverse)
library(infer)
library(broom)
library(openintro)
library(kableExtra)
library(tidytext)
library(lubridate)

library(praise)
set.seed(4747)
```


###  The Data

The data this week comes from [Harvard's Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=hdl:1902.1/12379) by way of [Mine Çetinkaya-Rundel, David Robinson, and Nicholas Goguen-Compagnoni](https://github.com/dgrtwo/unvotes/blob/7eb7034314ff79c49c9e0785fcd9d216fa04cf14/DESCRIPTION#L6).

Original Data citation:  
> Citation: Erik Voeten "Data and Analyses of Voting in the UN General Assembly" Routledge Handbook of International Organization, edited by Bob Reinalda (published May 27, 2013). Available at SSRN: http://ssrn.com/abstract=2111149



```{r}
unvotes <- read_csv("unvotes.csv")
roll_calls <- read_csv("roll_calls.csv")
issues <- read_csv("issues.csv")
```


```{r eval = FALSE, echo = FALSE}
library(unvotes)
data(unvotes)
data(roll_calls)
data(issues)
```

```{r}
issues %>%
  select(issue, short_name) %>%
  table()
```


```{r}
vote_issues <- unvotes %>%
  full_join(issues, by = "rcid") %>%
  full_join(roll_calls, by = "rcid")
  
```


## Voting patterns by date

```{r}
vote_issues %>%
  filter(country %in% c("China","United States", "Canada", "Taiwan", "India")) %>%
  mutate(year = year(date)) %>%
  filter(!is.na(issue)) %>%
  filter(year >= 1980) %>%
  group_by(year, country, short_name) %>%
  mutate(propyes = mean(vote == "yes"), propno = mean(vote == "no")) %>%
  ggplot(aes(x = year, y = propyes, color = country)) + 
  geom_point() + 
  geom_line() +
  facet_wrap(~issue)
```


```{r}
library(viridis)
vote_issues %>%
  filter(country %in% c("China","United States", "Canada", "Taiwan", "India")) %>%
  #mutate(country = ifelse(country == "Taiwan", "China", country)) %>%
  mutate(year = year(date)) %>%
  filter(!is.na(issue)) %>%
  #filter(year >= 1980) %>%
  group_by(year, country, short_name) %>%
  ggplot(aes(y = country, x = year, color = vote)) + 
  geom_jitter(height = .1) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~issue)
```  


```{r}
vote_issues %>%
  filter(country %in% c("China","United States", "Canada", "Taiwan", "India")) %>%
  mutate(year = year(date)) %>%
  filter(!is.na(issue)) %>%
  filter(year >= 1980) %>%
  filter(country == "United States") %>%
  #filter(short_name == "me") %>%
  ggplot(aes(x = year, fill = vote)) + 
  geom_area(stat = "bin", position = "fill") +
  scale_fill_viridis(discrete = TRUE) + 
  ylab("percentage") + 
  facet_wrap(~issue)

```

```{r}
praise()
```

