---
title: "Water Insecurity"
author: "Jo Hardin"
date: "1/28/2025"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## The Data

This week we're exploring water insecurity data featured in the article [Mapping water insecurity in R with tidycensus](https://waterdata.usgs.gov/blog/acs-maps/)!

> Water insecurity can be influenced by number of social vulnerability indicators—from demographic characteristics to living conditions and socioeconomic status —that vary spatially across the U.S. This blog shows how the tidycensus package for R can be used to access U.S. Census Bureau data, including the American Community Surveys, as featured in the “Unequal Access to Water ” data visualization from the USGS Vizlab. It offers reproducible code examples demonstrating use of tidycensus for easy exploration and visualization of social vulnerability indicators in the Western U.S.


The average household size was scraped from: https://worldpopulationreview.com/state-rankings/average-household-size-by-state

```{r}
water_insecurity_2022 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-01-28/water_insecurity_2022.csv') |> 
  select(-geometry)
water_insecurity_2023 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-01-28/water_insecurity_2023.csv') |> 
  select(-geometry)

ave_household <- readr::read_csv("ave_household.csv")

```

## Wrangling

Inspiration from @nrennie, https://github.com/nrennie/tidytuesday/tree/main/2025/2025-01-28.

We are looking to compare the percent of water insecurity in 2023 versus 2022, across states.
According to the GitHub repository, the population is number of people and the plumbing variable is number of households.
After summarizing the number of households over an entire state, we multiply by the average household size in that state (to get the number of people with water insecurity).

```{r}
water_insec <- water_insecurity_2022 |> 
  inner_join(water_insecurity_2023, by = "geoid", suffix = c(".2022", ".2023")) |> 
  tidyr::separate(col = name.2022, into = c("city", "state"), sep = ", ") |> 
  select(-name.2023, -percent_lacking_plumbing.2022, -percent_lacking_plumbing.2023)


water_insec  
```


```{r}
diff_plumb <- water_insec |> 
  group_by(state) |> 
  summarize(total_2022 = sum(total_pop.2022, na.rm = TRUE),
         total_2023 = sum(total_pop.2023, na.rm = TRUE),
         no_plumb_2022 = sum(plumbing.2022, na.rm = TRUE),
         no_plumb_2023 = sum(plumbing.2023, na.rm = TRUE)) |> 
  full_join(ave_household, by = "state") |> 
  mutate(perc_no_plumb_2022 = 100*no_plumb_2022 * size / total_2022,
         perc_no_plumb_2023 = 100*no_plumb_2023 * size / total_2023,
         change_plumbing = perc_no_plumb_2023 - perc_no_plumb_2022)
```

```{r}
data(state)
state_abbreviations <- data.frame(state.name, state.abb) |> 
  rbind(c("Puerto Rico", "PR")) |> 
  rbind(c("District of Columbia", "DC")) |> 
  mutate(state = as.factor(state.name)) 
```



## Plotting 

We plot both the difference (2023 minus 2022) in water insecurity as well as the proportion of the population with water insecurity in 2023.



```{r}
#| fig-cap: "For each state, we calculate the percentage of the population who is water insecure in both 2023 and 2022.  The x-axis gives the difference in percentages 2023 minus 2022. Alaska had an additional 0.5% of its population as water insecure in 2023 as compared to 2022. South Dakota had a decrease of 0.3% of its population as water insecure in 2023 as compared to 2022."
#| fig-alt: "Bar plot with change in percent of population as water insecure from 2022 to 2023. The x-axis gives the difference in percentages 2023 minus 2022. Alaska had an additional 0.5% of its population as water insecure in 2023 as compared to 2022. South Dakota had a decrease of 0.3% of its population as water insecure in 2023 as compared to 2022."

diff_plumb |> 
  inner_join(state_abbreviations, by = "state") |>
  mutate(nudge = ifelse(change_plumbing > 0, 0.05, -0.05)) |> 
  ggplot(aes(x = change_plumbing, y = fct_reorder(state.abb, change_plumbing))) + 
  geom_col(aes(fill = change_plumbing)) + 
  geom_text(aes(label = fct_reorder(state.abb, change_plumbing), 
                x = change_plumbing + nudge), 
                size = 3) + 
  theme_minimal() + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()) + 
  labs(x = "Change in percent of population lacking plumbing between 2022 and 2023",
       y = "",
       fill = "difference in insecurity\n 2023 minus 2022") + 
  scale_fill_continuous(low = "blue", high = "orange")
```


```{r}
#| fig-cap: "Note that while Alaska (approximately 1.3% of the population) and South Dakota (almost 0% of the population) are on the extremes of the percent of the population who is water insecure, the change from 2022 to 2023 does not monotonically match to the percentage of the state who is water insecure."
#| fig-alt: "Bar plot with percent of population as water insecure on the x axis and state on the y axis.  Alaska is the state with the highest percentage of the population as water insecure, at 1.3%. South Dakota is the state with the lowest percentage of the population as water insecure, at close to zero percent."
#| fig-format: png

diff_plumb |> 
  inner_join(state_abbreviations, by = "state") |>
  mutate(nudge = ifelse(change_plumbing > 0, 0.05, -0.05)) |> 
  ggplot(aes(x = perc_no_plumb_2023, y = fct_reorder(state.abb, change_plumbing))) + 
  geom_col(aes(fill = perc_no_plumb_2023)) + 
  geom_text(aes(label = fct_reorder(state.abb, change_plumbing), 
                x = perc_no_plumb_2023 + 0.05),
                size = 3) + 
  theme_minimal() + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()) + 
  labs(x = "Percent of population lacking plumbing in 2023",
       y = "",
       fill = "percent of insecurity 2023") + 
  scale_fill_continuous(low = "blue", high = "orange")
```


```{r}
praise()
```



