---
title: "Chess Matches"
author: "Jo Hardin"
date: "10/01/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## The Data

The chess dataset this week comes from Lichess.org via [Kaggle/Mitchell J](https://www.kaggle.com/datasets/datasnaek/chess/data).

> This is a set of just over 20,000 games collected from a selection of users on the site Lichess.org.

Use the data to explore -

1. What the common opening moves? By ranks?
2. How many turns does a game last based on player ranking?
3. What move patterns explain the game outcome?

Thank you to [Havisha Khurana](https://github.com/havishak) for curating this week's dataset.


## Collapsing the type of opening

There are a **lot** of opening moves.  In order to have some understanding of the patterns, we'll collapse the different opening moves into the base opening (e.g., before the `:`, `|`, or `#`).

```{r}
chess <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-01/chess.csv') |> 
  mutate(opening_main = case_when(
    str_detect(opening_name, ":") ~ str_extract(opening_name, ".+?(?=:)"),
    str_detect(opening_name, "\\|") ~ str_extract(opening_name, ".+?(?= \\|)"),
    str_detect(opening_name, "#") ~ str_extract(opening_name, ".+?(?= #)"),
    TRUE ~ opening_name)) |> 
  mutate(num_moves = str_count(moves, "\\S+"))
```


## Opening moves

Let's keep only the top 10 opening moves.

```{r}
top_open <- chess |> group_by(opening_main) |> 
  summarize(open_count = n()) |> 
  arrange(desc(open_count)) |> 
  filter(open_count > 500) |> 
  select(opening_main) |> 
  pull()

top_open

chess_top <- chess |> 
  filter(opening_main %in% top_open)
```


```{r}
chess_top |> 
  ggplot(aes(x = opening_main, fill = winner)) + 
  geom_bar(position = "fill") + 
  #theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)) + 
  scale_fill_brewer(palette = "Dark2")  + 
  coord_flip() + 
  labs(y = "", x = "",
       title = "Opening moves")
```

To look at the differences of opening strategies as a function of the player ratings, we'll use only the top 5 opening strategies.

```{r}

top_open_5 <- chess |> group_by(opening_main) |> 
  summarize(open_count = n()) |> 
  arrange(desc(open_count)) |> 
  filter(open_count > 500) |> 
  select(opening_main) |> 
  pull()

top_open_5

chess |> 
  filter(opening_main %in% top_open_5) |> 
  ggplot(aes(x = white_rating, y = black_rating, color = opening_main)) + 
  geom_point() +
  scale_color_brewer(palette = "Spectral")
```


## Model

Just for fun, let's run a logistic regression.  Seems like there isn't really much of a relationship between the ratings, the number of moves, whether or not the game/players were rated and the out come of whether the Sicilian Defense was used.

```{r}
chess <- chess |> 
  mutate(sicilian = ifelse(opening_main == "Sicilian Defense", 1, 0))

glm(sicilian ~ black_rating + white_rating + num_moves + rated, 
    data = chess, family = "binomial") |> 
  broom::augment(type.predict = "response") 

glm(sicilian ~ black_rating + white_rating + num_moves + rated, 
    data = chess, family = "binomial") |> 
  broom::augment(type.predict = "response") |> 
  ggplot(aes(x = .fitted, y = sicilian)) + 
  geom_point(alpha = 0.1)
```



```{r}
praise()
```


