---
title: "Dr. Who"
author: "Jo Hardin"
date: "11/28/2023"
format: html
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(tidytext)
library(praise)
library(paletteer)
library(ggforce)
library(networkD3)
library(plotly)
```



# The Data

The data this week comes from Wikipedia's [List of Doctor Who episodes](https://en.wikipedia.org/wiki/List_of_Doctor_Who_episodes_(2005%E2%80%93present) via the [{datardis} package](https://cran.r-project.org/package=datardis) by [Jonathan Kitt](https://github.com/KittJonathan/datardis).

```{r}
directors <- read_csv("drwho_directors.csv")
episodes <- read_csv("drwho_episodes.csv")
writers <- read_csv("drwho_writers.csv")
```



```{r}
#| fig-alt: Scatterplot of rating by number of views in the UK, faceted by the writer of the episode.  All of the scatterplots show a moderately postive correlation.  Gareth Roberts and Stephen Moffat have high rated and highly watched episodes; Chris Chibnall has low rated and low watched episodes.

p <- episodes |>
  left_join(writers, by = "story_number") |>
  group_by(writer) |>
  mutate(n_episodes = n()) |>
  filter(n_episodes >= 5) |>
  ggplot(aes(x = uk_viewers, y = rating, color = writer, label = episode_title)) + 
  geom_point() +
  facet_wrap( ~ writer) + 
  ggthemes::scale_color_colorblind() +
  guides(color="none") + 
  labs(x = "number of views, UK (millions)", y = "appreciation index",
       title = "Evaluting Dr. Who writers")

ggplotly(p, tooltip = c("episode_title"))
```

It isn't clear where the rating came from.  Maybe IMDb? Let's scrape the IMDb website to see if the ratings are the same as what was used in the original dataset. From what I know, British ratings are provided by randomly sampled households using a device in their house to provide a measure of how good the show is.  It is called the Appreciation Index.

```{r}
library(rvest)
library(purrr)

scrape_imdb <- function(i){
  imdb_rating <- paste0("https://www.imdb.com/title/tt0436992/episodes/?season=", i, sep = "") |>
  read_html() |>
  html_elements("span[aria-label^='IMDb']") |>
  html_text2() |>
  str_extract("[\\d\\.]+(?=/)") |>
  as.numeric() 
  
  episode <- paste0("https://www.imdb.com/title/tt0436992/episodes/?season=", i, sep = "") |>
  read_html() |>
  html_nodes(xpath = "//div[@class='ipc-title__text']") |>
  html_text2() |>
  str_extract("[S\\d\\.E\\d]+(?=\\s)") |>
    data.frame() |>
    setNames("ep.seas") |>
  mutate(season_number = as.numeric(str_extract(ep.seas, "\\d+")),
         episode_number = as.numeric(str_extract(ep.seas, "(?<=E)\\d+"))) 
  
  data.frame(imdb_rating, episode)
    
}

imdb_who <- 1:13 |>
  map_dfr(scrape_imdb)

```


```{r}
#| fig-alt: Scatter plot comparing the IMDb ratings (which were scraped from IMDb) and the ratings in the original dataset.  The values are have moderately strong positive correlation but clearly are not identical.

episodes |>
  left_join(imdb_who, by = c("season_number", "episode_number")) |>
  ggplot(aes(x = imdb_rating, y = rating)) + 
  geom_point() + 
  labs(x = "IMDb rating", y = "Appreciation Index",
       title = "Comparing two different ratings")
```


```{r}
praise()
```

