---
title: "Science Foundation Ireland Grants Commitments"
author: "Jo Hardin"
date: February 24, 2026
format: 
  html:
    embed-resources: true
    code-fold: true
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## The Data

This week we're exploring Irish STEM research and ancillary projects funded by Science
Foundation Ireland (SFI) since its foundation in 2000 to its dissolution on 31st
July 2024. The data comes from [Ireland's Open Data Portal](https://data.gov.ie/dataset/science-foundation-ireland-grants-commitments)

> SFI was the national foundation in Ireland for investment in scientific and
> engineering research. Consequently, SFI invested in those academic
> researchers and research teams who were most likely to generate new
> knowledge, leading edge technologies and competitive enterprises in the
> fields of science, technology, engineering and maths (STEM).



```{r}
sfi_grants <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-02-24/sfi_grants.csv') |> 
  mutate(AI = str_detect(str_to_lower(proposal_title),
                      "\\bai\\b|\\ba[.]i[.]\\b|\\bartificial\\s+intelligence\\b|(neural net)|(machine learn)|(deep learn)")) |> 
  mutate(year = year(start_date)) |> 
  drop_na(proposal_title) |> 
  filter(year <= 2024)
```



```{r}
ai_data <- sfi_grants |> 
  filter(AI)
```



```{r}
#| fig.cap: The proportion of grants given in a particular year across the space of grants.  That is, over 20% of the AI grants were given in 2023. Fewer than 2% of the non-AI grants were given in 2009.
#| fig.alt: A line graph with year on the x-axis and proportion of the total grants given per year on the y-axis. There are two lines, one for AI grants and one for non-AI grants. The AI grants have become a higher proportion of total grants in recent years. The non-AI grants have been steady for the last decade or so.

sfi_grants |> 
  mutate(AI = ifelse(AI, "AI grant", "non-AI grant")) |> 
  group_by(AI, year) |> 
  summarize(count = n(), average = mean(current_total_commitment)) |> 
  mutate(count_total = sum(count), prop_total = count / count_total) |> 
  ggplot(aes(x = year, y = prop_total, color = AI)) + 
  geom_line() + 
  geom_point(aes(size = average)) + 
  theme_minimal() + 
  labs(x = "",
       y = "",
       title = "Proportion of grants given",
       color = "type of grant",
       size = "average grant\n(in Euros)") +
  scale_color_manual(values = c("AI grant" = "#648FFF", "non-AI grant" = "#FE6100"))
```




```{r}
#| echo: false
#| include: false

p <- sfi_grants |> 
  mutate(AI = ifelse(AI, "AI grant", "non-AI grant")) |> 
  group_by(AI, year) |> 
  summarize(count = n(), average = mean(current_total_commitment)) |> 
  mutate(count_total = sum(count), prop_total = count / count_total) |> 
  ggplot(aes(x = year, y = prop_total, color = AI)) + 
  geom_line() + 
  geom_point(aes(size = average)) + 
  theme_minimal() + 
  labs(x = "",
       y = "",
       title = "Proportion of grants given",
       color = "type of grant",
       size = "average grant\n(in Euros)") +
  scale_color_manual(values = c("AI grant" = "#648FFF", "non-AI grant" = "#FE6100"))


png(filename = "sfi.png", width = 550, height = 450)

p

dev.off()

```


```{r}
praise()
```

