---
title: "Solar Eclipse"
author: "Jo Hardin"
date: "04/09/2024"
format: html
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## The Data

This week we're looking at the paths of solar eclipses in the United States. The data comes from [NASA's Scientific Visualization Studio](https://svs.gsfc.nasa.gov/5073).


```{r}
eclipse_annular_2023 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-04-09/eclipse_annular_2023.csv')
eclipse_total_2024 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-04-09/eclipse_total_2024.csv')
eclipse_partial_2023 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-04-09/eclipse_partial_2023.csv')
eclipse_partial_2024 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-04-09/eclipse_partial_2024.csv')

```


Putting the 2023 datasets together and the 2024 datasets together so that they can be plotted on the same map.  The `time` variable represents the total duration of the eclipse: the amount of time from when the moon first comes in contact with the sun to when the moon is no longer in contact with the sun.


```{r}
tot_23 <- eclipse_annular_2023 |>
  mutate(time = eclipse_6 - eclipse_1,
         eclipse = "total") |>
  select(state, name, lat, lon, time, eclipse)

part_23 <- eclipse_partial_2023 |>
  mutate(time = eclipse_5 - eclipse_1,
         eclipse = "partial") |>
  select(state, name, lat, lon, time, eclipse)

eclipse_23 <- rbind(tot_23, part_23) |>
  mutate(time = as.numeric(time))
  
```

```{r}
tot_24 <- eclipse_total_2024 |>
  mutate(time = eclipse_6 - eclipse_1,
         eclipse = "total") |>
  select(state, name, lat, lon, time, eclipse)

part_24 <- eclipse_partial_2024 |>
  mutate(time = eclipse_5 - eclipse_1,
         eclipse = "partial") |>
  select(state, name, lat, lon, time, eclipse)

eclipse_24 <- rbind(tot_24, part_24) |>
  mutate(time = as.numeric(time))
  
```




```{r}
#| fig-cap: In blue, the path of the 2023 annular solar eclipse.  
#|   The times represent the difference (in seconds) between when 
#|   the moon first contacts the sun and when the moon last contacts 
#|   the sun.
#| fig-alt: A map of the United States where points are plotted for
#|   each city. Each point is colored based on the amount of time
#|   the eclipse is happening. That is, the total time the moon 
#|   contacts the sun at all. The duration of the eclipse depends on
#|   both the longitude (basically the distance from the path of 
#|   the full eclipse) and the latitude (the distance from the 
#|   equator).


library(ggnewscale)
states <- map_data("state")

ggplot(states) +
  geom_polygon(fill = "white", color = "black", 
               aes(long, lat, group=group)) +
  geom_point(data = filter(eclipse_23, lat < 51 & lat > 24 & lon < 0 & eclipse == "total"), 
             aes(x = lon, y = lat, color = time), size = .1) +
  labs(x = "", y = "") + 
  scale_colour_gradientn(colors = c("blue", "turquoise")) +
  labs(color = "time of overlap (sec)\nannular eclipse") + 
  new_scale_color() + 
  geom_point(data = filter(eclipse_23, lat < 51 & lat > 24 & lon < 0 & eclipse == "partial"), 
             aes(x = lon, y = lat, color = time), size = .001) +
  scale_colour_gradientn(colors = c("red", "pink")) +
  ggtitle("2023 annular solar eclipse") +
  labs(color = "time of overlap (sec)\npartial eclipse")
```



```{r}
#| fig-cap: In blue, the path of the 2024 total solar eclipse.  The 
#|   times represent the difference (in seconds) between when the 
#|   moon first contacts the sun and when the moon last contacts the sun.
#| fig-alt: A map of the United States where points are plotted for
#|   each city. Each point is colored based on the amount of time
#|   the eclipse is happening. That is, the total time the moon 
#|   contacts the sun at all. The duration of the eclipse depends on
#|   both the longitude (basically the distance from the path of 
#|   the full eclipse) and the latitude (the distance from the 
#|   equator).

ggplot(states) +
  geom_polygon(fill = "white", color = "black", 
               aes(long, lat, group=group)) +
  geom_point(data = filter(eclipse_24, lat < 51 & lat > 24 & lon < 0 & eclipse == "total"), 
             aes(x = lon, y = lat, color = time), size = .0001) +
  labs(x = "", y = "") + 
  scale_colour_gradientn(colors = c("blue", "turquoise")) +
  labs(color = "time of overlap (sec)\ntotal eclipse") + 
  new_scale_color() + 
  geom_point(data = filter(eclipse_24, lat < 51 & lat > 24 & lon < 0 & eclipse == "partial"), 
             aes(x = lon, y = lat, color = time), size = .0001) +
  scale_colour_gradientn(colors = c("red", "pink")) +
  ggtitle("2024 total solar eclipse") +
  labs(color = "time of overlap (sec)\npartial eclipse")
```



```{r}
praise()
```


