---
title: "Democracy & Dictatorship"
author: "Jo Hardin"
date: "11/05/2024"
format: html
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## The Data

It's Election Day in the United States. If you are able to do so, vote!

To celebrate, we're looking at a dataset about Democracy and Dictatorship.

> This dataset updates [pacl](https://xmarquez.github.io/democracyData/reference/pacl.html) 
> with more countries and coverage from 1950 to 2020, as described in C. 
> Bjørnskov and M. Rode. "Regime types and regime change: A new dataset on 
< democracy, coups, and political institutions". In: *The Review of International Organizations*
> 15.2 (2020), pp. 531-551. [DOI: 10.1007/s11558-019-09345-1](https://link.springer.com/article/10.1007/s11558-019-09345-1).
> The full data and codebook can be downloaded [here](http://www.christianbjoernskov.com/bjoernskovrodedata/).


```{r}
democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-05/democracy_data.csv') |> 
  mutate(regime = case_when(
    stringr::str_detect(regime_category, "(?i)colony") ~ "Colony",
    regime_category == "Civilian dictatorship" ~ "Civilian dictatorship",
    regime_category == "Parliamentary democracy" ~ "Parliamentary democracy",
    regime_category == "Military dictatorship" ~ "Military dictatorship",
    regime_category == "Presidential democracy" ~ "Presidential democracy",
    regime_category == "Mixed democratic" ~ "Mixed democratic",
    regime_category == "Royal dictatorship" ~ "Royal dictatorship",
    regime_category == "Part of the USSR" ~ "Part of the USSR",
    TRUE ~ "other"
  )) |> 
  mutate(country_name = case_when(
    country_name == "United States" ~ "USA",
    country_name == "United Kingdom" ~ "UK",
    country_name == "Gambia, The" ~ "Gambia",
    country_name == "Congo, Dem. Rep." ~ "Democratic Republic of the Congo",
    country_name == "Congo, Republic of" ~ "Republic of Congo",
    #country_name == "" ~ "",
    #country_name == "" ~ "",
    #country_name == "" ~ "",
    TRUE ~ country_name
  )) |> 
  mutate(regime = factor(regime, 
                            levels = c("Royal dictatorship",
                                       "Military dictatorship",
                                       "Civilian dictatorship",
                                       "Colony",
                                       "Part of the USSR",
                                       "Presidential democracy",
                                       "Mixed democratic",
                                       "Parliamentary democracy", "other")))

```


```{r}

world <- map_data("world")

world |> 
  mutate(the = str_detect(region, "(?i)congo")) |> 
  dplyr::filter(the) |> 
  select(region, the) |> 
  distinct()

democracy_data |> 
  mutate(the = str_detect(country_name, "(?i)congo")) |> 
  dplyr::filter(the) |> 
  select(country_name, the) |> 
  distinct()

```


```{r}
democracy_data |> 
  select(regime_category) |> 
  group_by(regime_category) |> 
  summarize(count = n()) |> 
  arrange(desc(count))
```


```{r}
ghana <- democracy_data |> 
  dplyr::filter(country_code == "GHA")

usa <- democracy_data |> 
  dplyr::filter(country_code == "USA")

```


```{r}
ghana |> 
  ggplot(aes(x = year, y = regime_category, color = regime_category)) + 
  geom_point()

usa |> 
  ggplot(aes(x = year, y = regime_category, color = regime_category)) + 
  geom_point()
```



## Mapping the data

```{r}
world <- map_data("world")

map_data <- democracy_data |> 
  right_join(world, by = c("country_name" = "region"))
```


### To figure out which countries don't match:

```{r}
democracy_data |> 
  anti_join(world, by = c("country_name" = "region")) |> 
  select(country_name) |> 
  distinct()

world |> 
  anti_join(democracy_data, by = c("region" = "country_name")) |> 
  select(region) |> 
  distinct()
```

```{r}
map_data2 <- map_data |> 
  dplyr::filter(year >= 2000)
```

```{r}
#| fig-cap: A map describing the political regimes over time.  Unfortunately, I can't get the map to transition smoothly, so it is a pretty terrible graph.
#| fig-alt: A map of the world where each country is colored by its type of political regime. The map transitions over time to see how regime type changes over time.  

library(plotly)
p <- map_data2 |> 
  #dplyr::filter(year == 2020) |> 
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = regime, frame = year)) + 
  theme_void()

animation_opts(
  p,
  frame = 500,
  transition = 0,
  easing = "linear",
  redraw = FALSE,
  mode = "next"
)

ggplotly(p)
```




```{r}
praise()
```





