---
title: "Economic Diversity and Student Outcomes"
author: "Jo Hardin"
date: "09/10/2024"
format: html
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```


## Data

The data explores economic diversity and student outcomes. The dataset this week comes from [Opportunity Insights](https://opportunityinsights.org/data/) and is described in its [codebook](https://opportunityinsights.org/wp-content/uploads/2023/07/CollegeAdmissions_Codebook.pdf).

Unfortunately, the data aren't easy to interpret.  In the work below, we use the variable `attend`, but we aren't sure what it means.  It is vaguely close to the proportion of students [out of those who take the SAT/ACT and are in a given income bracket] who would attend a given college.

```{r}
college_admissions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-10/college_admissions.csv')
```

```{r}
college_admissions |> 
  select(name) |> 
  distinct()
```

There are 139 institutions in the dataset.


Next, we try to suss out the variable `attend`, we would have expected it to sum to 1 given a particular income group. However, it doesn't sum to one for 

```{r}
college_admissions |> 
  group_by(par_income_bin) |> 
  summarize(prob_attend_sum = sum(attend, na.rm = TRUE)) |> 
  ggplot(aes(x = par_income_bin, y = prob_attend_sum)) +
  geom_point() + 
  labs(x = "income bracket of parents (percentile)",
       y = "sum of `attend` variable over the income bracket",
       title = "unsuccessfully trying to understand the `attend` variable")
```

Regardless, we will continue with the `attend` variable as some measure of attendance as a function of income percentile of the parents.

```{r}
#| fig-cap: By hovering over the lines we can see the relationship between the parental income percentile and the attendance variable for each of the 139 institutions in the study. The private schools all tend to go **up** for the highest incomes. The majority of public schools go **down** for the highest incomes.
#| fig-alt: Line plot with parental income percentile on the x-axis and attendance on the y-axis.  Each line represents a different institution. The graph is faceted by whether the school is a public or private school.  Within the private schools, we see that attendance goes up for the highest incomes.  Within the public schools, we see that attendance generally goes down for the highest incomes.

library(plotly)
library(crosstalk)

admiss <- SharedData$new(college_admissions, ~name)
gg1 <- admiss |> 
  ggplot(aes(x = par_income_bin, y = attend, color = public,
             text = paste0("school: ", name))) + 
  geom_point() + 
  geom_line() + 
  facet_grid(~ public) + 
  labs(color = "Public School",
       x = "parental income bin",
       y = "something about % attendance")


gg2 <- ggplotly(gg1, tooltip = "text")

plotly::highlight(gg2, on = "plotly_hover")
```



```{r}
praise()
```



