---
title: "MarioKart"
author: "Jo Hardin"
date: "5/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width=10, eval.after = 'fig.cap')
library(tidyverse)
library(infer)
library(broom)
library(tidytext)
library(lubridate)

library(praise)
set.seed(4747)
```

### The Data


The data this week comes from [Mario Kart World Records](https://mkwrs.com/) and contains world records for the classic (if you're a 90’s kid) racing game on the Nintendo 64.

```{r}
mariokart <- read_csv("records.csv")
```

### Change in records

By looking at the world record vs time (for the `Three Lap` shortcut races), it seems there are some pretty easily identifiable jumps in time.

```{r}
mariokart %>%
  filter(type == "Three Lap")  %>%
  filter(shortcut == "Yes") %>%
  ggplot(aes(x = date, y = time)) + 
  geom_point() + 
  geom_step() + 
  facet_wrap(~track)
```

But what if we wanted to measure the percent decrease in time?  The new variable of interest will be:

$$\% \mbox{ decrease} = \Bigg(\frac{\mbox{time}_i - \mbox{time}_{i+1}}{\mbox{time}_i}\Bigg)\cdot100 = \Bigg(1 - \frac{\mbox{time}_{i+1}}{\mbox{time}_i}\Bigg)\cdot100$$
It is interesting to note that the biggest percent reductions in time happen at the very beginning, except for Choco Mountain.


```{r fig.height=9}
mariokart %>%
  filter(type == "Three Lap", shortcut == "Yes") %>%
  group_by(track) %>%
  mutate(time_lag = dplyr::lag(time)) %>%
  mutate(perc_decr = (1 - time / time_lag)*100) %>%
  ggplot(aes(x = date)) +
  geom_point(aes( y = perc_decr)) +
  geom_step(aes( y = perc_decr, color = "Percent decrease")) +
  geom_step(aes( y = time/3, color = "Record time")) +
  scale_y_continuous(sec.axis = sec_axis(~.*3, name = "Time (sec)")) +
  scale_color_manual(values = c("#D89A9E", "#114B5F")) +
  facet_wrap(~track) +
  labs(
    y = "Decrease [%]",
    x = "Date",
    color = "Trendline:",
#    strip.text.y = element_text(angle = 0, size = rel(4)),
    caption = "Tidy Tuesday Plot: @hardin47 | Data: MarioKart") +
    ggtitle("MarioKart record times") +
  theme(legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        plot.title = element_text(size = 18))
```





```{r}
praise()
```