---
title: "Repair Cafe"
author: "Jo Hardin"
date: April 7, 2026
format: 
  html:
    embed-resources: true
    code-fold: true
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse) # ggplot, lubridate, dplyr, stringr, readr...
library(praise)
```

# Repair Cafes Worldwide

The dataset this week comes from the [Repair Monitor](https://dashboard.repairmonitor.org/?language=en), which has been compiling data from Repair Cafes worldwide since 2015.
Repair Cafe branches bring together volunteer fixers to help people learn how to repair household items that are broken. 

Note: There appears to be some uncertainty (by submitters to the source data) of what to put in `repair_info_source` and `repair_info_url`. We included the questions for these fields to aid in the interpretation of the data.

> As carbon-hungry consumer production and its subsequent waste surge to all-time highs, experts say that the concept can help curb pollution while promoting a more circular economy. 

- What kinds of items are most easily repaired?
- What are the most common reasons that items can't be repaired? 
- Which countries have seen the most growth in Repair Cafe branches?
- Is GenAI becoming more popular than YouTube as a source of useful information for repairers?

Thank you to [Jen Richmond](https://github.com/jenrichmond) for curating this week's dataset.


# The Data

```{r}
repairs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-04-07/repairs.csv')
repairs_text <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-04-07/repairs_text.csv')
```



```{r}
repairs |>
  group_by(brand) |>
  summarize(count = n()) |>
  arrange(desc(count))
```


```{r}
repairs |>
  filter(brand %in% c("Philips", "Bosch", "Sony", "Miele", "Tefal")) |>
  group_by(brand) |>
  summarize(mean_repair = mean(repairability, na.rm = TRUE))
```
```{r}
repairs |>
  filter(brand %in% c("Philips", "Bosch", "Sony", "Miele", "Tefal")) |>
  ggplot() +
  geom_jitter(aes(x = repairability, y = brand), alpha = .1)+
  geom_point(aes(x = repairability, y = brand), color = "red")
```


```{r}
repairs |>
  filter(brand %in% c("Philips", "Bosch", "Sony", "Miele", "Tefal")) |>
  filter(estimated_year_of_production > 1910, estimated_year_of_production < 2026) |> 
  ggplot() +
  geom_jitter(aes(x = estimated_year_of_production, y = repairability,
                 color = brand), width = 0, alpha = 0.4) +
  geom_point(aes(x = estimated_year_of_production, y = repairability, color = brand))
```



```{r}
#| fig.cap: "For the 5 most frequent brands in the Repair Cafe dataset, how do the production year and the repairability compare? (Repairability is on a scale of 1 (difficult) to 10 (easy).) There do not seem to be any strong trends across time or brand."
#| fig.alt: "Box plots of the year of production compared to the repairability. Repairability is on a scale of 1 (difficult) to 10 (easy). The boxplots are also broken down by the brand, using the following brands: Bosch, Miele, Philips, Sony, and Tefal. There do not appear to be any trends linking the production year and the repairability score."

repairs |>
  filter(brand %in% c("Philips", "Bosch", "Sony", "Miele", "Tefal")) |>
  filter(estimated_year_of_production > 1910, estimated_year_of_production < 2026) |> 
  filter(!is.na(repairability)) |> 
  filter(repairability > 0) |>
  ggplot() +
  geom_boxplot(aes(y = estimated_year_of_production, x = as.factor(repairability),
                 fill = brand), color = "black") + 
  geom_jitter(aes(y = estimated_year_of_production, x = as.factor(repairability)), 
              alpha = 0.1, size = 0.2) + 
  geom_hline(yintercept = 2015) + 
  theme_minimal() + 
  labs(x = "repairability", y = "production year",
       title = "Is it harder to repair older items?")
```


```{r}
#| echo: false
#| include: false

p <- repairs |>
  filter(brand %in% c("Philips", "Bosch", "Sony", "Miele", "Tefal")) |>
  filter(estimated_year_of_production > 1910, estimated_year_of_production < 2026) |> 
  filter(!is.na(repairability)) |> 
  ggplot() +
  geom_boxplot(aes(y = estimated_year_of_production, x = as.factor(repairability),
                 fill = brand), color = "black") + 
  geom_jitter(aes(y = estimated_year_of_production, x = as.factor(repairability)), 
              alpha = 0.1, size = 0.2) + 
  geom_hline(yintercept = 2015) + 
  theme_minimal() + 
  labs(x = "repairability", y = "production year",
       title = "Is it harder to repair older items?")


png(filename = "repair.png", width = 650, height = 450)

p

dev.off()

```

```{r}
praise()
```



