---
title: "Art and Artists"
author: "Jo Hardin"
date: "1/12/2021"
output: html_document
---

The data today come from the [Tate Art Museum](https://github.com/tategallery/collection).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(infer)
library(broom)
library(countrycode)
library(praise)
set.seed(4747)
```


## Input data


```{r}
artists <- readr::read_csv("artists.csv")
artwork <- readr::read_csv("artwork.csv")
```


Note, over half of the pieces in the collection are by Joseph Turner.  We might be curious to analyze the data without his pieces.

```{r}
artwork %>%
  group_by(artistId) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  head()

artists %>%
  filter(id == 558) 
```

Let's combine the datasets so that if we want to use gender to ask questions about the art, the information is available.  Note there are 6 pieces of art without a connection to the artists database.  And there are 194 artists without art in the dataset.

```{r}
artartist1 <- inner_join(artists, artwork, by = c("id" = "artistId"))
artartist2 <- full_join(artists, artwork, by = c("id" = "artistId"))

nrow(artwork)
nrow(artartist1)
nrow(artartist2)
```

### Location

Both `placeOfBirth` and `placeOfDeath` are given by "City, Country", and it might be easier to visualize or model if we had only country.

```{r}
artists <- artists %>%
  separate(placeOfBirth, c("cityBirth", "countryBirth") ,sep = ", ", fill = "left") %>%
  separate(placeOfDeath, c("cityDeath", "countryDeath") ,sep = ", ", fill = "left") 

artists %>% group_by(countryBirth) %>%
  mutate(count = n()) %>%
  filter(count >= 20) %>%
  ungroup() %>%
  ggplot(aes(x = countryBirth, y = yearOfBirth)) + 
  geom_boxplot()

```


#### Viz artwork


```{r}
artartist1 %>%
  filter(id != 558) %>%
  ggplot(aes(x = year, color = gender, fill = gender)) + 
  geom_histogram()

```

```{r}
praise()
```











